---
title: "DIGIT : Questionnaires and Surveys: Analyses with R"
author: "David Plans"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  bookdown::html_document2:
link-citations: yes

---


**Preparation and session set up**

Before turning to the tasks below, please install the packages by running the code below this paragraph. If you have already installed the packages mentioned below, then you can skip ahead ignore this section. To install the necessary packages, simply run the following code - it may take some time (between 1 and 5 minutes to install all of the libraries so you do not need to worry if it takes some time).

```{r prep_01, echo=T, eval = F, message=FALSE, warning=FALSE}
# install packages
install.packages("knitr")
install.packages("lattice")
install.packages("tidyverse")
install.packages("likert")
install.packages("MASS")
install.packages("psych")
install.packages("viridis")
install.packages("ggplot2")
install.packages("here")
install.packages("flextable")
```

You can now activate the packages by running the code chunk below. 

```{r prep_02, message=FALSE, warning=FALSE}
# set options
options(stringsAsFactors = F)         # no automatic data transformation
options("scipen" = 100, "digits" = 4) # suppress math annotation
# install packages
# 
# DIGIT TODO : look carefully at the packages loaded, as they will give you a clue as to what functions from what packages you might use to accomplish the tasks below

library(knitr)
library(lattice)
library(tidyverse)
library(likert)
library(MASS)
library(psych)
library(viridis)
library(ggplot2)
library(here)
library(flextable)

```

Once you have installed R, RStudio, and have also initiated the session by executing the code shown above, you are good to go. Look carefully at the packages loaded above, as they will give you a clue as to what functions from what packages you might use to accomplish the tasks below


## Line graphs for Likert-scaled data{-}

DIGIT TODO: load a data set (ldat) which contains Likert-scaled variables. This data set represents fictitious rating of students from courses about how satisfied they were with their learning experience. The response to the Likert item is numeric so that *strongly disagree/very dissatisfied* would get the lowest (1) and *strongly agree/very satisfied* the highest numeric value (5).

```{r line_0a1, message=FALSE, warning=FALSE}
# define color vectors so you can use them in charts if you want
clrs1 <- c("firebrick4",  "gray70", "darkblue")
clrs2 <- c("firebrick4", "firebrick1", "gray70", "blue", "darkblue")

# load the exercise dataset
ldat <- base::readRDS(url("https://slcladal.github.io/data/lid.rda", "rb"))
```

DIGIT TODO inspect the `ldat` data set and create a table with the first 10 rows of the data as a dataframe (using as.data.frame())

```{r line_01b, echo = F}
# try to use tidyverse pipelines such as:
ldat %>%

# insert your code here to complete expression above (fill out the pipeline)
```

DIGIT TODO : The ldat data set has only two columns: a column labeled *Course* which has three levels (*German*, *Japanese*, and *Chinese*) and a column labeled *Satisfaction* which contains values from 1 to 5 which represent values ranging from *very dissatisfied* to *very satisfied*. Now that you have data resembling a Likert-scaled item from a questionnaire, please will display the data in a cumulative line graph using ggplot.

```{r line_03, echo=T, message=FALSE, warning=FALSE}
# create cumulative density plot
ldat %>%

#insert your code here.
  
```


The satisfaction of the German course was the lowest as the red line shows the highest density (frequency of responses) of *very dissatisfied* and *dissatisfied* ratings. The students in our fictitious data set were most satisfied with the Chinese course as the blue line is the lowest for *very dissatisfied* and "dissatisfied" ratings while the difference between the courses shrinks for "satisfied" and *very satisfied*. The Japanese language course is in-between the German and the Chinese course.  



Most commonly, the data for visualization comes from tables of absolute frequencies associated with a categorical or nominal variable. The default way to visualize such frequency tables are pie charts and bar plots. Modify the data we've got so far and create a table then a bar chart.


DIGIT TODO : modify the data to get counts and percentages. 

```{r pie_01, echo = T, eval=T, message=FALSE, warning=FALSE}
# create bar plot data. Perhaps use dyplyr's group_by, summarise and mutate functions in order to get counts and percentages
bdat <- ldat %>%

#insert your code here.
  
```


DIGIT TODO inspect the new data set and create an overview table, again using the bdat dataframe

```{r pie_02, echo = F}
bdat %>%
  as.data.frame() %>%

```

## Bar plot

DIGIT TODO : Create a bar plot of bdat to display frequency information across categorical variable levels using ggplot.

```{r bar_01, echo = T, results='hide', message=FALSE, warning=FALSE}
# bar plot
bdat %>%

#insert code here

```


```{r fin}
sessionInfo()
```




